### 2024-06-02 2024-5-13/sections/introduction.tex
fix the syntax error in introduction

\medskip
\noindent 
\emph{Contribution.} In this work, we focus on the class of string constraints with regular membership and string-length function, where the counting operators may occur (called RECL for brevity). We make the following contributions.
\begin{itemize}
  \item We propose an automata-theoretical approach for solving RECL constraints.
  Our main idea is to represent the counting operators by cost registers in cost-enriched finite automata (CEFA, see Section~\ref{sec:automaton} for the definition), instead of unfolding them explicitly. The string-length function is modeled by cost registers as well. The satisfiability of RECL constraints is reduced to the nonemptiness problem of CEFA w.r.t. a linear integer arithmetic (LIA) formula. According to the results from~\cite{atva2020}, an LIA formula can be computed to represent the potential values of registers in CEFA.
Thus, the nonemptiness of CEFA w.r.t. LIA formulas can be reduced to the satisfiability of LIA formulas, which is then tackled by off-the-shelf SMT solvers. Moreover, we propose techniques to reduce the sizes (i.e. the number of states and transitions) of CEFA, in order to achieve better performance.
%
  \item For complex RECL constraints where the counting operators may be nested with the other counting operators or complement operators, we propose optimization techniques to improve the performance. Furthermore, for the satisfiable RECL constraints where the solutions of the LIA formulas can be obtained, but the models of the original RECL constraint cannot be generated,  we propose optimization techniques to accelerate the model generation process. 
%
%  \item 
%  Combined with the size-reduction techniques mentioned above, the register representation of regex-counting and string-length in CEFA entails an efficient algorithm for solving RECL constraints. 
 %
\item We implement the algorithm on top of OSTRICH, resulting in a string solver called $\ostrichrecl$. Finally, we utilize a collection of benchmark suites comprising 49,843 problem instances in total to evaluate the performance of $\ostrichrecl$. The experiment results show that $\ostrichrecl$ solves the RECL constraints more efficiently than the state-of-the-art string solvers (see Table~\ref{tab:table_overall_eval}-\ref{tab:results_model-gen}). For instance, $\ostrichrecl$ solves 49,638 out of 49,843 instances, while the best state-of-the-art solvers solve only 47,721 instances. We also do experiments to justify some of our technical choices and evaluate the performance of the optimizations. 
%
% 1,000 benchmark instances where the counting and length bounds are large or the counting operators are nested with some other counting operators or complement operators, $\ostrichrecl$ solves at least $278$ more instances than the other solvers, while spending only half or less time per instance on average. 
\end{itemize}

\medskip
\noindent
\emph{Related work.} 
We discuss more related work on regexes with counting, string-length function, and automata with registers/counters.  
%
Determinism of regexes with counting was investigated in~\cite{GGM12,CL15}. 
%
Real-world regexes in programming languages include features beyond classical regexes, e.g., the greedy/lazy Kleene star, capturing groups, and back references. Real-world regexes have been addressed in symbolic execution of Javascript programs~\cite{LMK19} and string constraint solving~\cite{CF+22}. 
Nevertheless, the counting operators are still unfolded explicitly in~\cite{CF+22}. 
%
The Trau tool focuses on string constraints involving flat regular languages and string-length function and solves them by computing LIA formulas that define the Parikh images of flat regular languages~\cite{z3trau}. The Slent tool solves the string constraints involving string-length function by encoding them into so-called length-encoded automata, then utilizing symbolic model checkers to solve their nonemptiness problem~\cite{WC+18}. However, neither Trau nor Slent supports counting operators explicitly, in other words, counting operators in regexes should be unfolded before solved by them.
%
Cost registers in CEFAs are different from registers in (symbolic) register automata~\cite{ra,sra}:  In register automata, registers are used to store input values and  can only be compared for equality/inequality, while in CEFAs, cost registers are used to store integer-valued costs and can be updated by adding/subtracting integer constants and constrained by the accepting conditions which are LIA formulas.
%
Therefore, cost registers in CEFAs are more like counters in counter automata/machines~\cite{Minsky67}, that is, CEFAs can be obtained from counter machines by removing  transition guards and adding accepting conditions. 
%
%For the automata model, the definition of registers of CEFAs resembles the counters of the counter automata class~\cite{Minsky67}, but CEFAs contain no guards in each transition. 
%
Counting-set automata were proposed in~\cite{redos_lenka,HS+23} to quickly match a subclass of regexes with counting. Moreover, a variant of nondeterministic counter automata, called bit vector automata, was proposed recently in \cite{GKM23} to enable fast matching of a more expressive class of regexes with counting.   Nevertheless, the nonemptiness problem of these automata was not considered, and it is unclear whether these automata models can be used for solving string constraints with regex-counting and string-length.

@@

A preliminary version of this paper has appeared in the proceedings of SETTA 2023~\cite{Denghang2023}. This paper extends the SETTA paper in the following three aspects.
\begin{itemize}
\item To improve the performance of complex RECL constraints where the counting operators are nested with some other counting operators or complement operators, various optimizations are proposed. 
\item To deal with the issue that the satisfiability of LIA formulas can be solved, but models of the original RECL constraints cannot be generated, we propose techniques to optimize the model generation process.  
\item The benchmarks are extended to cove the situations that the counting operators are nested with some other counting operators or complement operators, and the experiments are extended accordingly.  
\end{itemize}

>>>>

\medskip
\noindent 
\emph{Contribution.} In this work, we focus on the class of string constraints with regular membership and string-length function, where the counting operators may occur (called RECL for brevity). We make the following contributions.
\begin{itemize}
  \item We propose an automata-theoretical approach for solving RECL constraints.
 Our main idea is to represent the counting operators by cost registers in cost-enriched finite automata (CEFA, see Section~\ref{sec:automaton} for the definition), instead of unfolding them explicitly. The string-length function is modeled by cost registers as well. The satisfiability of RECL constraints is reduced to the non-emptiness problem of CEFA w.r.t. a linear integer arithmetic (LIA) formula. According to the results from~\cite{atva2020}, an LIA formula can be computed to represent the potential values of registers in CEFA.
Thus, the non-emptiness of CEFA w.r.t. LIA formulas can be reduced to the satisfiability of LIA formulas, which is then tackled by off-the-shelf SMT solvers. Moreover, we propose techniques to reduce the sizes (i.e., the number of states and transitions) of CEFA in order to achieve better performance.
%
  \item For complex RECL constraints where the counting operators may be nested with the other counting operators or complement operators, we propose optimization techniques to improve the performance. Furthermore, for the satisfiable RECL constraints where the solutions of the LIA formulas can be obtained, but the models of the original RECL constraint cannot be generated,  we propose optimization techniques to accelerate the model generation process. 
%
%  \item 
%  Combined with the size-reduction techniques mentioned above, the register representation of regex-counting and string-length in CEFA entails an efficient algorithm for solving RECL constraints. 
 %
\item We implement the algorithm on top of OSTRICH, resulting in a string solver called $\ostrichrecl$. Finally, we utilize a collection of benchmark suites comprising 49,843 problem instances in total to evaluate the performance of $\ostrichrecl$. The experiment results show that $\ostrichrecl$ solves the RECL constraints more efficiently than the state-of-the-art string solvers (see Table~\ref{tab:table_overall_eval}-\ref{tab:results_model-gen}). For instance, $\ostrichrecl$ solves 49,638 out of 49,843 instances, while the best state-of-the-art solvers solve only 47,721 instances. We also do experiments to justify some of our technical choices and evaluate the performance of the optimizations. 
%
% 1,000 benchmark instances where the counting and length bounds are large or the counting operators are nested with some other counting operators or complement operators, $\ostrichrecl$ solves at least $278$ more instances than the other solvers, while spending only half or less time per instance on average. 
\end{itemize}

\medskip
\noindent
\emph{Related work.} 
We discuss more related work on regexes with counting, string-length function, and automata with registers/counters.  
%
Determinism of regexes with counting was investigated in~\cite{GGM12,CL15}. 
%
Real-world regexes in programming languages include features beyond classical regexes, e.g., the greedy/lazy Kleene star, capturing groups, and back references. Real-world regexes have been addressed in symbolic execution of Javascript programs~\cite{LMK19} and string constraint solving~\cite{CF+22}. 
Nevertheless, the counting operators are still unfolded explicitly in~\cite{CF+22}. 
%
The Trau tool focuses on string constraints involving flat regular languages and string-length function and solves them by computing LIA formulas that define the Parikh images of flat regular languages~\cite{z3trau}. The Slent tool solves the string constraints involving string-length function by encoding them into so-called length-encoded automata, then utilizing symbolic model checkers to solve their non-emptiness problem~\cite{WC+18}. However, neither Trau nor Slent supports counting operators explicitly, in other words, counting operators in regexes should be unfolded before solved by them.
%
Cost registers in CEFAs are different from registers in (symbolic) register automata~\cite{ra,sra}:  In register automata, registers are used to store input values and can only be compared for equality/inequality, while in CEFAs, cost registers are used to store integer-valued costs and can be updated by adding/subtracting integer constants and constrained by the accepting conditions which are LIA formulas.
%
Therefore, cost registers in CEFAs are more like counters in counter automata/machines~\cite{Minsky67}, that is, CEFAs can be obtained from counter machines by removing transition guards and adding accepting conditions. 
%
%For the automata model, the definition of registers of CEFAs resembles the counters of the counter automata class~\cite{Minsky67}, but CEFAs contain no guards in each transition. 
%
Counting-set automata were proposed in~\cite{redos_lenka,HS+23} to quickly match a subclass of regexes with counting. Moreover, a variant of nondeterministic counter automata, called bit vector automata, was proposed recently in \cite{GKM23} to enable fast matching of a more expressive class of regexes with counting.   Nevertheless, the non-emptiness problem of these automata was not considered, and it is unclear whether these automata models can be used for solving string constraints with regex-counting and string-length.

@@

A preliminary version of this paper has appeared in the proceedings of SETTA 2023~\cite{Denghang2023}. This paper extends the SETTA paper in the following three aspects.
\begin{itemize}
\item To improve the performance of complex RECL constraints where the counting operators are nested with some other counting operators or complement operators, various optimizations are proposed. 
\item To deal with the issue that the satisfiability of LIA formulas can be solved, but models of the original RECL constraints cannot be generated, we propose techniques to optimize the model generation process.  
\item The benchmarks are extended to cover the situations where the counting operators are nested with some other counting operators or complement operators, and the experiments are extended accordingly.  
\end{itemize}
